### 2023-07-04 paper/sec-analysis.tex
Update sec-analysis.tex

The expected impact of the FPF data on global PDF fits is assessed by a profiling procedure~\cite{Paukkunen:2014zia, Schmidt:2018hvu, AbdulKhalek:2018rok, HERAFitterdevelopersTeam:2015cre}, based on minimizing the function
%TODO correlations are also included here, as we hopefully get there in the end -- if not, remove terms involving correlations+pseudodata
\begin{equation}
\chi^2 = 
\sum_{i=1}^{N_{\textrm{bins}}} 
\frac{\left(  \sigma_i^{\textrm{pd}}
            + \Gamma_i^{\alpha,\textrm{pd}}
              b_\alpha^{\textrm{pd}}
            - \sigma_i^{\textrm{th}}
            - \Gamma_i^{\beta,\textrm{th}}
              b_\beta^{\textrm{th}}
     \right)^2
     }{\Delta_i^2}
+ \sum_\alpha (b_\alpha^{\textrm{pd}})^2
+ \sum_\beta  (b_\beta^{\textrm{th}})^2 \, .
\label{eq:profilingchi2}
\end{equation}
Here the pseudodata 
$\sigma_i^{\textrm{pd}}$ 
is obtained from the central theoretical prediction 
$\sigma_i^{\textrm{th}}$ 
for each bin $i$ out of $N_{\textrm{bins}}$ by varying it within bounds obtained from the statistical (systematic) uncertainties 
$\delta_i^{\textrm{stat}}$ ($\delta_i^{\textrm{syst}}$), 
obtained in the procedure described in Section~\ref{sec:pseudo-data_generation}, 
as

\begin{equation}
\sigma_i^{\textrm{pd}}
=
\sigma_i^{\textrm{th}}
\left( 1 + r_i \sqrt{     (\delta_i^{\textrm{stat}})^2
                      + (f \delta_i^{\textrm{syst}})^2 }
\right).
\end{equation}
Here, $r_i$ is a univariate Gaussian random number and $f = 0.5$ an effective correction factor 
accounting for the improved constraining power of the same pseudodata set using correlated systematics, compared to solely adding all sources of uncertainty in quadrature.
%
The correlated uncertainties for the pseudodata and the theoretical prediction 
are contained in the nuisance parameter vectors $b^{\textrm{pd}}$ and $b^{\textrm{th}}$, respectively, and the uncorrelated uncertainties in $\Delta_i$.
%
Their effect on $\sigma^{\textrm{th}}$ and $\sigma^{\textrm{pd}}$
is described by the matrices $\Gamma_i^{\textrm{pd}}$ and $\Gamma_i^{\textrm{th}}$.
The indices $\alpha$ and $\beta$ then run over the uncertainty nuisance parameters for the pseudodata and the theoretical prediction, respectively.
%
The nuisance parameter values $b^{\textrm{th(min)}}$ that minimize Eq.~\eqref{eq:profilingchi2} give the central PDFs $f'_0$ optimized to the profiled dataset in the form
\begin{equation}
f_0' = f_0
      + \sum_\beta b_\beta^{\textrm{th(min)}} 
        \left(  \frac{f_\beta^+   -  f_\beta^- }{2}
              -    b_\beta^{\textrm{th(min)}}
                \frac{f_\beta^+ + f_\beta^- - 2f_0}{2}
        \right),
\end{equation}
where $f_0$ is the original central PDF and the up and down variation eigenvectors are given by $f^+, f^-$. The decrease in the uncertainties of the shifted PDFs indicates the enhancement that including the FPF data in the global PDF fit could bring. Here, the profiling studies are performed using version 2.2.1 of the \textsc{xFitter} open-source QCD analysis framework~\cite{Alekhin:2014irh, Bertone:2017tig, xFitter:2022zjb, xFitter:web}.
%

To this end, a new interface between  {\sc\small PineAPPL} and {\sc\small xFitter} has been developed and is available with the open-source {\sc\small xFitter} code.

>>>>

The profiling method applied to Hessian PDF fits is described in~\cite{Paukkunen:2014zia, Schmidt:2018hvu, AbdulKhalek:2018rok, HERAFitterdevelopersTeam:2015cre} and is based
on minimizing a goodness-of-fit error function defined as
\begin{equation}
\chi^2 = 
\sum_{i=1}^{N_{\textrm{bin}}} 
\frac{\left(  \mathcal{O}_i^{\rm (exp)}
            + \Gamma_i^{\alpha,\textrm{pd}}
              b_\alpha^{\textrm{(pd)}}
            - \mathcal{O}_i^{\textrm{th}}
            - \Gamma_i^{\beta,\textrm{th}}
              b_\beta^{(\textrm{th})}
     \right)^2
     }{\delta_{{\rm stat},i}^2}
+ \sum_\alpha \lp b_\alpha^{(\textrm{pd})}\rp^2
+ \sum_\beta  \lp b_\beta^{\textrm{(th)}}\rp ^2 \, .
\label{eq:profilingchi2}
\end{equation}
The pseudodata 
$\mathcal{O}_i^{\rm (exp)}$ 
is obtained from the central theoretical prediction 
$\mathcal{O}_i^{(\textrm{th})}$
for each bin $i$ out of $N_{\textrm{bin}}$ by fluctuating
it within its uncertainties, following either
Eq.~(\ref{eq:pseudo_data}), when accounting for the correlations
between systematic uncertainties, or Eq.~(\ref{eq:pseudo_data_v2}),
in the case in which errors are added in quadrature.

The correlated uncertainties for the pseudodata and for the theoretical prediction 
are contained in the nuisance parameter vectors $b^{(\textrm{pd})}$ and $b^{(\textrm{th})}$, respectively, and the total uncorrelated uncertainties are denoted by $\delta_{{\rm stat},i}$.
%
The effect of the nuisance parameters
on the observables $\mathcal{O}_i^{\rm (exp)}$ and $\mathcal{O}_i^{\rm (th)}$
is described by the matrices $\Gamma_i^{\textrm{pd}}$ and $\Gamma_i^{\textrm{th}}$.
%
The indices $\alpha$ and $\beta$ then run over the uncertainty nuisance parameters for the pseudodata and the theoretical prediction, respectively.
%
The nuisance parameter values $b_\beta^{\textrm{(th,min)}}$ that minimize Eq.~\eqref{eq:profilingchi2} give the central PDFs $f'_0$ optimized to the profiled dataset in the form
\begin{equation}
f_0' = f_0
      + \sum_\beta b_\beta^{\textrm{(th,min)}} 
        \left(  \frac{f_\beta^+   -  f_\beta^- }{2}
              -    b_\beta^{\textrm{(th,min)}}
                \frac{f_\beta^+ + f_\beta^- - 2f_0}{2}
        \right),
\end{equation}
where $f_0$ is the original central PDF and the up and down variation eigenvectors are given by $f^+, f^-$.
%
The reduction in the uncertainties of the profiled PDFs indicate the impact
of the projected data with respect to the assumed prior PDF set.

The profiling studies carried out in this work are performed using v2.2.1 of the \textsc{xFitter} open-source QCD analysis framework~\cite{Alekhin:2014irh, Bertone:2017tig, xFitter:2022zjb, xFitter:web}.
%
To this end, a new interface between  {\sc\small PineAPPL} and {\sc\small xFitter} has been developed and is available in {\sc\small xFitter}.
%
All the experimental and theoretical data files used in the analysis, including
the  {\sc\small PineAPPL}  grids, are available
from the {\sc\small xFitter} repository.
